Analysis of the Anomalous Doppler Effect from Quantum Theory to Classical Dynamics Simulations

### 2025-05-22 paper modification/CPB-tex-template/CPB Template-LaTeX/paper_v2.tex
add new

\begin{abstract}
A quantum model incorporating angular momentum conservation is developed to analyze the Normal and Anomalous Doppler Effects, demonstrating that the resonance condition is strongly influenced by the angular momentum of the wave. The resonance condition involving wave angular momentum is examined numerically, and the energy exchange ratio between the electron's parallel and gyrokinetic motion during resonance with the electromagnetic wave is simulated, exhibiting strong agreement with quantum theoretical predictions.
\end{abstract}

\textbf{Keywords: Anomalous Doppler Effect, Resonant condition, Angular monmentum Conservation } %no more than four sets of keywords should be provided

\textbf{PACS:} %no more than four PACS codes should be provided: check https://cpb.iphy.ac.cn/UserFiles/File/PACS2010Regular-Edition.pdf

@@

In 1967, Artsimovich \cite{artsimovich1967thermal} reported discrepancies in tokamak experiments: the electron temperature estimated from diamagnetic signals was significantly higher than that derived from electrical conductivity measurements. Although unrecognized at the time, this anomaly may represent the first experimental observation of ADE. In 1968, B. B. Kadomtsev \cite{kadomtsev1968electric} identified ADE as the underlying mechanism, in which electrons undergo velocity scattering from the longitudinal to the transverse direction under resonant conditions. This process enhances the diamagnetic effect beyond what would be expected from thermal motion alone. Subsequently, a range of ADE-related phenomena have been observed, including electron beam scattering in magnetic field vacuum tubes \cite{frank1960optics}, wave radiation \cite{spong2018first,liu2019intense,gorozhanin1997waves}, and runaway electron instabilities in tokamaks \cite{sajjad2007runaway,castejon2003particle}. Practical applications based on ADE have also emerged, such as high-power microwave generation and runaway electron suppression in tokamaks \cite{guo2018control,zhang2024self}.

The physics of the Anomalous Doppler Effect (ADE) was first elucidated through quantum analysis by Frank and Ginzburg \cite{frank1960optics,ginzburg1979nonlinear}. In this work, we extend Ginzburg's quantum framework by incorporating  the conservation of angular momentum to provide a more rigorous analysis of ADE. This approach yields new insights into the relationship between wave angular momentum and ADE under resonant conditions, which it is referred to Anomalous Doppler Resonance (ADR). For an electron moving in a magnetic field and interacting with an external electromagnetic (EM) wave, the general resonance condition is given by:
\(
    \omega = m\omega_{ce} + \vec{k} \cdot \vec{v},
\)
where $\vec{k}$ is the wave vector, $\omega_{ce}$ is the electron cyclotron frequency (here $\omega_{ce}>0$ ), $\vec{v}$ is the electron velocity, and $\omega$ is the wave angular frequency, $m = 0, \pm1, \pm2, \pm3, \ldots$ represent the Landau level\cite{coppi1976slide}. Specially,  for plane EM waves, we find that resonance is restricted to the fundamental harmonics ($m = \pm1$) due to spin angular momentum conservation, reducing the condition to
\(
    \omega = \pm\omega_{ce} + \vec{k} \cdot \vec{v}.
\)
where the negative sign refers to ADR condition while the positive sign refers to Normal Doppler Resonance (NDR) condition. This represents a significant constraint compared to previous theoretical treatments\cite{dendy1987classical} which suggested possible resonance at all harmonic orders ($m = \pm1, \pm2, \dots$).
Despite the  simplicity of the  model, our analysis demonstrates that angular momentum conservation plays a crucial role in EM wave-electron resonance - an aspect that, to the best of our knowledge, has not been previously addressed in the literature.

@@

The induced linearly polarized wave along \( \vec{B}_0 \) can be decomposed into a combination of a right-hand circularly polarized wave (\( m = 1 \)) and a left-hand circularly polarized wave (\( m = -1 \)), such that
\(
\vec{E}_w = \vec{E}_R + \vec{E}_L,
\)
where
\(
\vec{E}_R = \frac{1}{2} E_0 (\vec{e}_x + i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)],
\vec{E}_L = \frac{1}{2} E_0 (\vec{e}_x - i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)].
\)
%If the wavevector \( \vec{k} \) lies in the \( y \)-\( z \) plane with a crossing angle \( \theta_k \) relative to the \( z \)-axis, then the new coordinate unit vectors for the wave field expression should be rotated accordingly to align with the direction of \( \vec{k} \). The transformed basis vectors are:
%\begin{equation}
%\begin{pmatrix}
%\vec{e}_x' \\
%\vec{e}_y'
%\end{pmatrix}
%=
%\begin{pmatrix}
%1 & 0 & 0 \\
%0 & \cos\theta_k & \sin\theta_k
%\end{pmatrix}
%\cdot
%\begin{pmatrix}
%\vec{e}_x \\
%\vec{e}_y \\
%\vec{e}_z
%\end{pmatrix}
%\end{equation}
The magnetic field of E.M wave is
\begin{equation}
\vec{B}_w = \frac{\vec{k} \times \vec{E}_w}{\omega}
\end{equation}

The six-dimensional phase space of an electron, described by its position $\bm{r}$ and momentum $\bm{p}$, are presented in the equations below. The vectors $\bm{E}$ and $\bm{B}$ represent the total field, including both static and electromagnetic components. Here, $c$ denotes the speed of light in vacuum, $e$ represents the electron’s charge and $m_0$ is the electron’s mass in the rest frame.

@@

To simulate the evolution of $\bm{r}$ and $\bm{p}$, the above system is discretized using the Volume-Preserving Algorithm. Let $j$ denote the iteration step and $\text{Cay}(\bm{A})$ represent the Cayley transform of matrix $\bm{A}$:

@@

The evolution of the electron’s motion is shown in Fig.~\ref{fig:kinetic_evolution}. As the electron accelerates from stationary in the electrostatic field (Fig.~\ref{fig:kinetic_evolution}(b)), the resonant frequencies increase simultaneously (Fig.~\ref{fig:kinetic_evolution}(a)). The change of parallel velocity caused by electromagnetic wave can be quantified as \( \Delta v = v_z - v_{zE0} \) as shown in Fig.~\ref{fig:kinetic_evolution}(c), where \( v_z \) represents the parallel velocity under the given scenario, while \( v_{zE0} \) denotes the parallel velocity resulting solely from the electrostatic field, which can be calculated using a theoretical equation as

@@

The cyclotron velocity is shown in Fig.~\ref{fig:kinetic_evolution}(d). The work done by electromagnetic wave is shown in fig. 5(e), which can be calculated by integrating the power with time as 
\(
E_{\parallel \text{emw}} = \int P_{\parallel \text{emw}} dt ,
\)
and 
\(
P_{\parallel \text{emw}} = -e \vec{v}_\perp \times \vec{B}_{\perp \text{emw}} \cdot v_z.
\)
Since all discrete data points are available from the simulation, it is not difficult to integrate all the discrete data over time. Fig.~\ref{fig:kinetic_evolution}(f) shows the gyro-kinetic energy evolution with time, where 
\(
E_\perp = \frac{1}{2} m_e v_\perp^2.
\)

@@

The Anomalous Doppler Effect begins to emerge when the time reaches \( 113 \tau_{ce} \), where \( \omega_{\text{ADE}} = \omega \) as shown in Fig.~\ref{fig:kinetic_evolution}(a). At this point, the parallel velocity begins to scatter into the perpendicular direction, evident from the decrease in \( \Delta v_z \) and the increase in \( v_{\perp} \) as seen in Fig.~\ref{fig:kinetic_evolution}(c) and Fig.~\ref{fig:kinetic_evolution}(d). During the resonant period, the changes in parallel kinetic and gyro-kinetic energies caused by the electromagnetic wave are calculated as \( \Delta T_{21} = \Delta E_{||\text{emw}} \approx -96.2 \times 10^{-5} \, \text{eV} \) and \( \Delta E_{\perp} \approx 39.3 \times 10^{-5} \, \text{eV} \). The energy transfer ratio is \( \frac{\Delta U_{21}}{\Delta T_{21}} \approx -0.408 \). According to quantum theory, the change ratio of \( \Delta U_{21}/\Delta T_{21} = -\hbar \omega_{ce}/\hbar \vec{k} \cdot \vec{v} = -0.3908 \), where \( \omega_{ce} \approx 3.51 \times 10^9 \, \text{s}^{-1} \), and \( k = 10^5 \, \text{m}^{-1} \), \( v_z = 90 \, \text{km/s} \). The quantum theory results are in good agreement with the numerical calculations. The energy change ratio is also derived in the Appendix, based on classical theory.
\subsection{Validation of the relationship with wave angular momentum. }
\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{Figure6.eps}% Here is how to import EPS art
\caption{Velocity evolution caused by induced wave with linear, right-circular and left-circular polarization. (a) The cyclotron velocity $v_\perp$ .(b) The change of parallel velocity caused by the electromagnetic wave.}
    \label{fig:AngularMomentum}
\end{figure}

@@

The three types of polarization waves are investigated under the same scenario as before. As a result, the right-hand circularly polarized wave (\( m = 1 \)) causes a velocity change only at around \( 23\tau_{ce} \), while the left-hand circularly polarized wave (\( m = -1 \)) causes a velocity change only at around \( 113\tau_{ce} \). This indicates that the right-circularly polarized wave is responsible for the NDE, while the left-hand circularly polarized wave is responsible for the ADE, which agrees well with the quantum analysis.

%The process can be understood as follows: For an electromagnetic wave with right-hand polarization propagating along the magnetic field, the electron in the magnetic field undergoes right-hand circular motion. When its parallel velocity satisfies the condition
%\(
%\omega - \vec{k} \cdot \vec{v} = \omega_{ce},
%\)
%known as the Normal Doppler Resonance (NDR) condition, the electron, in its co-moving cyclotron frame, perceives the wave frequency as equal to its rotational frequency. Consequently, the electron resonates and absorbs the electromagnetic wave, as indicated in Fig.~\ref{fig:AngularMomentum}(c) at \( 23\tau_{ce} \). According to the conservation of angular momentum and parallel momentum, both the cyclotron velocity and parallel velocity increase, as the electromagnetic wave carries positive angular momentum (in the same direction as the cyclotron electron's angular momentum) and parallel momentum, which correspond to \( \hbar \) and \( \hbar k \) in quantum physics.
%
%For a left-circularly polarized electromagnetic wave, the resonance and scattering process occurs when the electron velocity satisfies the condition
%\(
%\omega - \vec{k} \cdot \vec{v} = -\omega_{ce},
%\)
%known as the ADR. In the frame of the cyclotron electron, the electromagnetic wave has the same frequency and rotational direction as the electron, since the electron's velocity exceeds the wave phase velocity. Because the electromagnetic wave performs negative work on the electron, as shown in Fig.~\ref{fig:AngularMomentum}(c) at \( 113\tau_{ce} \), where \( E_{\text{emw}} \) is negative for a left-hand polarization wave, this is equivalent to the electron emitting an electromagnetic wave with the same properties as the induced wave. Since the emitted wave has left-hand circular polarization and positive momentum---corresponding to \( -\hbar \) and \( \hbar k \) in quantum physics---the cyclotron velocity increases while the parallel velocity decreases, to conserve angular momentum and momentum.
%
%For the left-circularly polarized wave, where the angular momentum \( m = -1 \), resonance occurs only at
%\(
%\omega = \vec{k} \cdot \vec{v} - \omega_{ce},
%\)
%as shown in Fig.~\ref{fig:7}. This behavior differs from previous results, where resonance for a plane electromagnetic wave could occur at any integer \( m \) satisfying
%\(
%\vec{k} \cdot \vec{v} + m\omega_{ce} - \omega = 0,
%\)
%as shown in Eq.~(36) and Eq.~(37) of Ref.~\cite{dendy1987classical}, but agrees well with angular momentum conservation analysis.
For a left-circularly polarized (LCP) electromagnetic wave, The angular momentum selection rule ($m = -1$) restricts resonance to only occur when:
\(
    \omega = \vec{k} \cdot \vec{v} - \omega_{ce},
\)
as confirmed numerically in Fig.~\ref{fig:7}. This represents a significant departure from previous classical treatments \cite{dendy1987classical} (Eqs. 36-37) that permitted resonance at arbitrary integer harmonics $m$, while being fully consistent with quantum angular momentum conservation principles.

@@

\section{Discussion }\label{sec:Discussion}
Based on the momentum and angular momentum conservation analysis, we analyze the case where $\vec{k}$ is oriented opposite to $v_{\parallel}$(or $\vec{B_0}$). In this case, if a cyclotron electron emits a photon with left-hand circular polarization and momentum $-\hbar\vec{k}$, where the angular momentum carried by photon is $\hbar$, then after the emission, the change in internal energy is $\Delta U = -\hbar\omega_{ce}<0$, and the change in translational kinetic energy $\Delta T = \hbar k v_{\parallel}>0$. However, if the emitting photon have right-circular polarization and momentum $-\hbar\vec{k}$, the change of internal energy becomes $\Delta U = \hbar\omega_{ce}>0$, while the translational kinetic energy still is $\Delta T = \hbar k v_{\parallel}>0$. This would violate the conservation of energy, as it is not possible for an electron to emit a photon while simultaneously increasing its total energy. Consequently, for a plane electromagnetic wave, only the left-circularly polarized component can resonate with an electron moving opposite to $v_{\parallel}$ (or $\vec{B_0}$). 
%Based on momentum and angular momentum conservation analysis, we examine the case where $\vec{k}$ is oriented opposite to $\vec{v}_{\parallel}$ (or $\vec{B_0}$). 
%\begin{itemize}
%    \item \textbf{Left-hand circular polarization case}: When a cyclotron electron emits a photon with:
%
%    \begin{itemize}
%        \item Left-hand circular polarization ($\sigma^-$)
%        \item Momentum $-\hbar\vec{k}$
%        \item Angular momentum $+\hbar$
%    \end{itemize}
%    the resulting energy changes are:
%    \begin{align}
%        \Delta U &= -\hbar\omega_{ce} < 0 \quad \text{(decrease in internal energy)} \\
%        \Delta T &= +\hbar k v_{\parallel} > 0 \quad \text{(increase in translational energy)}
%    \end{align}
%
%    \item \textbf{Right-hand circular polarization case}: If the emitted photon had:
%    \begin{itemize}
%        \item Right-hand polarization ($\sigma^+$)
%        \item Momentum $-\hbar\vec{k}$
%        \item Angular momentum $-\hbar$
%    \end{itemize}
%    the energy changes would instead be:
%    \begin{align}
%        \Delta U &= +\hbar\omega_{ce} > 0 \quad \text{(increase in internal energy)} \\
%        \Delta T &= +\hbar k v_{\parallel} > 0 \quad \text{(increase in translational energy)}
%    \end{align}
%\end{itemize}

The right-hand polarization case would violate energy conservation, as it requires the electron to \emph{emit} a photon while simultaneously \emph{gaining} total energy ($\Delta U + \Delta T > 0$). 
Therefore, for a plane electromagnetic wave, only the left-circularly polarized component can resonantly interact with an electron moving antiparallel to $\vec{v}_{\parallel}$ while satisfying all conservation laws.

\section{Conclusion}\label{sec:Conclusion}
This paper presents a simple yet useful method to analyze the resonant process of NDE and ADE. The quantum method, combined with an angular momentum conservation analysis, illustrates that the parameter $m$ in the resonant condition $\omega = \vec{k} \cdot \vec{v} + m\omega_{ce}$ is directly related to the angular momentum of the resonant wave. Numerical simulations based on the VPA method are also provided, confirming the correctness of the quantum results regarding both the angular momentum relationship between $m$ and the energy transfer ratio.

%\addcontentsline{toc}{chapter}{Appendix A: Appendix section heading}
\section*{Appendix A:Classical analysis of Anomalous Doppler Resonant}
here we would like give a brief derivation of energy transformation through classical dynamic equation:
\begin{align}
m_e \frac{d\vec{v}_\parallel}{dt} &= -e(\vec{v}_\perp \times \vec{B}_\perp) \label{eq:19} \\
m_e \frac{d\vec{v}_\perp}{dt} &= -e(\vec{v}_\perp \times \vec{B}_0 + \vec{v}_\parallel \times \vec{B}_\perp + \vec{v}_\perp \times \vec{B}_0) \label{eq:20}
\end{align}

Consider $\vec{B}_\perp = \frac{\vec{e}_k \times \vec{E}_\perp}{v_p}$, where $\vec{e}_k$ is the unit vector of wave vector of E.M. wave, which is along the $z$-axis. Using $\vec{v}_\parallel$ and $\vec{v}_\perp$ to dot both sides of \eqref{eq:19} and \eqref{eq:20}, substitute $\vec{B}_\perp$ and simplify the equations, we have

>>>>

\begin{abstract}
A quantum model incorporating angular momentum conservation is developed to analyze the Normal and Anomalous Doppler Effects, demonstrating that the resonance condition is strongly influenced by the angular momentum of the wave. The resonance condition involving wave angular momentum is examined numerically, and the energy exchange ratio between the electron's parallel and gyro-kinetic motion during resonance with the electromagnetic wave is simulated, exhibiting strong agreement with quantum theoretical predictions.
\end{abstract}

\textbf{Keywords: Anomalous Doppler Effect, Resonant condition, Angular momentum Conservation,  } %no more than four sets of keywords should be provided

\textbf{PACS: 33.35.+r, 02.60.Cb,  03.65.-w，12.20.-m} %no more than four PACS codes should be provided: check https://cpb.iphy.ac.cn/UserFiles/File/PACS2010Regular-Edition.pdf

@@

In 1967, Artsimovich \cite{artsimovich1967thermal} reported discrepancies in tokamak experiments: the electron temperature estimated from diamagnetic signals was significantly higher than that derived from electrical conductivity measurements. Although unrecognized at the time, this anomaly may represent the first experimental observation of ADE. In 1968, B. B. Kadomtsev \cite{kadomtsev1968electric} identified ADE as the underlying mechanism, in which electrons undergo velocity scattering from the longitudinal to the transverse direction under resonant conditions. This process enhances the diamagnetic effect beyond what would be expected from thermal motion alone. Subsequently, a range of ADE-related phenomena have been observed, including electron beam scattering in magnetic field vacuum tubes \cite{frank1960optics}, wave radiation \cite{spong2018first,liu2019intense,gorozhanin1997waves}, and runaway electron instabilities in tokamaks \cite{sajjad2007runaway,castejon2003particle}. The ADE has also given rise to practical applications, notably in high-power microwave generation and in mitigating runaway electrons in tokamak fusion reactors. \cite{guo2018control,zhang2024self}.

The physics of the Anomalous Doppler Effect (ADE) was first elucidated through quantum analysis by Frank and Ginzburg \cite{frank1960optics,ginzburg1979nonlinear}. In this work, we extend Ginzburg's quantum framework by incorporating  the conservation of angular momentum to provide a more rigorous analysis of ADE. This approach yields new insights into the relationship between wave angular momentum and ADE under resonant conditions, which it is referred to Anomalous Doppler Resonance (ADR). For an electron moving in a magnetic field and interacting with an external electromagnetic (EM) wave, the general resonance condition is given by:
\(
    \omega = m\omega_{ce} + \vec{k} \cdot \vec{v},
\)
where $\vec{k}$ is the wave vector, $\omega_{ce}$ is the electron cyclotron frequency (here $\omega_{ce}>0$ ), $\vec{v}$ is the electron velocity, and $\omega$ is the wave angular frequency, $m = 0, \pm1, \pm2, \pm3, \ldots$ represent the Landau level\cite{coppi1976slide}. Specially,  for plane EM waves, we find that resonance is restricted to the fundamental harmonics ($m = \pm1$) due to spin angular momentum conservation, reducing the condition to
\(
    \omega = \pm\omega_{ce} + \vec{k} \cdot \vec{v}
\)
,where the negative sign refers to ADR condition while the positive sign refers to Normal Doppler Resonance (NDR) condition. This represents a significant constraint compared to previous theoretical treatments\cite{dendy1987classical} which suggested possible resonance at all harmonic orders ($m = \pm1, \pm2, \dots$).
Despite the  simplicity of the  model, our analysis demonstrates that angular momentum conservation plays a crucial role in EM wave-electron resonance - an aspect that, to the best of our knowledge, has not been previously addressed in the literature.

@@

The induced linearly polarized wave along \( \vec{B}_0 \) can be decomposed into a combination of a right-hand circularly polarized wave (\( m = 1 \)) and a left-hand circularly polarized wave (\( m = -1 \)), such that
\(
\vec{E}_w = \vec{E}_R + \vec{E}_L,
\)
where
\(
\vec{E}_R = \frac{1}{2} E_0 (\vec{e}_x + i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)],
\vec{E}_L = \frac{1}{2} E_0 (\vec{e}_x - i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)].
\)
%If the wavevector \( \vec{k} \) lies in the \( y \)-\( z \) plane with a crossing angle \( \theta_k \) relative to the \( z \)-axis, then the new coordinate unit vectors for the wave field expression should be rotated accordingly to align with the direction of \( \vec{k} \). The transformed basis vectors are:
%\begin{equation}
%\begin{pmatrix}
%\vec{e}_x' \\
%\vec{e}_y'
%\end{pmatrix}
%=
%\begin{pmatrix}
%1 & 0 & 0 \\
%0 & \cos\theta_k & \sin\theta_k
%\end{pmatrix}
%\cdot
%\begin{pmatrix}
%\vec{e}_x \\
%\vec{e}_y \\
%\vec{e}_z
%\end{pmatrix}
%\end{equation}
The magnetic field of EM wave is
\begin{equation}
\vec{B}_w = \frac{\vec{k} \times \vec{E}_w}{\omega}
\end{equation}

The six-dimensional phase space of an electron, described by its position $\bm{r}$ and momentum $\bm{p}$, are presented in the equations below. The vectors $\bm{E}$ and $\bm{B}$ represent the total field, including both static and electromagnetic components. Here, $c$ denotes the speed of light in vacuum, $e$ represents the electron's charge and $m_0$ is the electron's mass in the rest frame.

@@

To simulate the evolution of $\bm{r}$ and $\bm{p}$, the above system is discretized using the Volume-Preserving Algorithm\cite{zhang2015volume,liu2016collisionless}. Let $j$ denote the iteration step and $\text{Cay}(\bm{A})$ represent the Cayley transform of matrix $\bm{A}$:

@@

The evolution of the electron's motion is shown in Fig.~\ref{fig:kinetic_evolution}. As the electron accelerates from stationary in the electrostatic field (Fig.~\ref{fig:kinetic_evolution}(b)), the resonant frequencies increase simultaneously (Fig.~\ref{fig:kinetic_evolution}(a)). The change of parallel velocity caused by electromagnetic wave can be quantified as \( \Delta v = v_z - v_{zE0} \) as shown in Fig.~\ref{fig:kinetic_evolution}(c), where \( v_z \) represents the parallel velocity under the given scenario, while \( v_{zE0} \) denotes the parallel velocity resulting solely from the electrostatic field, which can be calculated using a theoretical equation as

@@

The cyclotron velocity is shown in Fig.~\ref{fig:kinetic_evolution}(d). The work done by electromagnetic wave is shown in Fig.~\ref{fig:kinetic_evolution}(e), which can be calculated by integrating the power with time as 
\(
E_{\parallel \text{emw}} = \int P_{\parallel \text{emw}} dt ,
\)
and 
\(
P_{\parallel \text{emw}} = -e \vec{v}_\perp \times \vec{B}_{\perp \text{emw}} \cdot v_z.
\)
Since all discrete data points are available from the simulation, it is not difficult to integrate all the discrete data over time. Fig.~\ref{fig:kinetic_evolution}(f) shows the gyro-kinetic energy evolution with time, where 
\(
E_\perp = \frac{1}{2} m_e v_\perp^2.
\)

@@

The Anomalous Doppler Resonance begins to emerge when the time reaches \( 113 \tau_{ce} \), where \( \omega_{\text{ADE}} = \omega \) as shown in Fig.~\ref{fig:kinetic_evolution}(a). At this point, the parallel velocity begins to scatter into the perpendicular direction, evident from the decrease in \( \Delta v_z \) and the increase in \( v_{\perp} \) as seen in Fig.~\ref{fig:kinetic_evolution}(c) and Fig.~\ref{fig:kinetic_evolution}(d). During the resonant period, the changes in parallel kinetic and gyro-kinetic energies caused by the electromagnetic wave are calculated as \( \Delta T_{21} = \Delta E_{||\text{emw}} \approx -96.2 \times 10^{-5} \, \text{eV} \) and \( \Delta E_{\perp} \approx 39.3 \times 10^{-5} \, \text{eV} \). The energy transfer ratio is \( \frac{\Delta U_{21}}{\Delta T_{21}} \approx -0.408 \). According to quantum theory, the change ratio of \( \Delta U_{21}/\Delta T_{21} = -\hbar \omega_{ce}/\hbar \vec{k} \cdot \vec{v} = -0.3908 \), where \( \omega_{ce} \approx 3.51 \times 10^9 \, \text{s}^{-1} \), and \( k = 10^5 \, \text{m}^{-1} \), \( v_z = 90 \, \text{km/s} \). The quantum theory results are in good agreement with the numerical calculations. The energy change ratio is also derived in the Appendix, based on classical theory.
\subsection{Validation of the relationship with wave angular momentum. }
\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{Figure6.eps}% Here is how to import EPS art
\caption{Velocity evolution caused by induced wave with linear, right-circular and left-circular polarization. (a) The cyclotron velocity $v_\perp$ .(b) The change of parallel velocity caused by the electromagnetic wave.}
    \label{fig:AngularMomentum}
\end{figure}

@@

The three types of polarization waves are investigated under the same scenario as before. As a result, the right-hand circularly polarized (RHCP) wave (\( m = 1 \)) causes a velocity change only at around \( 23\tau_{ce} \), while the left-hand circularly polarized (LHCP) wave (\( m = -1 \)) causes a velocity change only at around \( 113\tau_{ce} \). This indicates that the RHCP wave is responsible for the NDE, while the LHCP wave is responsible for the ADE, which agrees well with the quantum analysis.

%The process can be understood as follows: For an electromagnetic wave with right-hand polarization propagating along the magnetic field, the electron in the magnetic field undergoes right-hand circular motion. When its parallel velocity satisfies the condition
%\(
%\omega - \vec{k} \cdot \vec{v} = \omega_{ce},
%\)
%known as the Normal Doppler Resonance (NDR) condition, the electron, in its co-moving cyclotron frame, perceives the wave frequency as equal to its rotational frequency. Consequently, the electron resonates and absorbs the electromagnetic wave, as indicated in Fig.~\ref{fig:AngularMomentum}(c) at \( 23\tau_{ce} \). According to the conservation of angular momentum and parallel momentum, both the cyclotron velocity and parallel velocity increase, as the electromagnetic wave carries positive angular momentum (in the same direction as the cyclotron electron's angular momentum) and parallel momentum, which correspond to \( \hbar \) and \( \hbar k \) in quantum physics.
%
%For a left-circularly polarized electromagnetic wave, the resonance and scattering process occurs when the electron velocity satisfies the condition
%\(
%\omega - \vec{k} \cdot \vec{v} = -\omega_{ce},
%\)
%known as the ADR. In the frame of the cyclotron electron, the electromagnetic wave has the same frequency and rotational direction as the electron, since the electron's velocity exceeds the wave phase velocity. Because the electromagnetic wave performs negative work on the electron, as shown in Fig.~\ref{fig:AngularMomentum}(c) at \( 113\tau_{ce} \), where \( E_{\text{emw}} \) is negative for a left-hand polarization wave, this is equivalent to the electron emitting an electromagnetic wave with the same properties as the induced wave. Since the emitted wave has left-hand circular polarization and positive momentum---corresponding to \( -\hbar \) and \( \hbar k \) in quantum physics---the cyclotron velocity increases while the parallel velocity decreases, to conserve angular momentum and momentum.
%
%For the left-circularly polarized wave, where the angular momentum \( m = -1 \), resonance occurs only at
%\(
%\omega = \vec{k} \cdot \vec{v} - \omega_{ce},
%\)
%as shown in Fig.~\ref{fig:7}. This behavior differs from previous results, where resonance for a plane electromagnetic wave could occur at any integer \( m \) satisfying
%\(
%\vec{k} \cdot \vec{v} + m\omega_{ce} - \omega = 0,
%\)
%as shown in Eq.~(36) and Eq.~(37) of Ref.~\cite{dendy1987classical}, but agrees well with angular momentum conservation analysis.
For a LHCP electromagnetic wave, The angular momentum selection rule ($m = -1$) restricts resonance to only occur when:
\(
    \omega = \vec{k} \cdot \vec{v} - \omega_{ce},
\)
as confirmed numerically in Fig.~\ref{fig:7}. This represents a significant departure from previous classical treatments \cite{dendy1987classical} (Eqs. 36-37) that permitted resonance at arbitrary integer harmonics $m$, while being fully consistent with quantum angular momentum conservation principles.

@@

\section{Discussion }\label{sec:Discussion}
Based on the momentum and angular momentum conservation analysis, Let's analyze the case where $\vec{k}$ is oriented opposite to $v_{\parallel}$(or $\vec{B_0}$). In this case, if a cyclotron electron emits a photon with left-hand circular polarization and momentum $-\hbar\vec{k}$, where the angular momentum carried by photon is $\hbar$, then after the emission, the change in internal energy is $\Delta U = -\hbar\omega_{ce}<0$, and the change in translational kinetic energy $\Delta T = \hbar k v_{\parallel}>0$. However, if the emitting photon have right-circular polarization and momentum $-\hbar\vec{k}$, the change of internal energy becomes $\Delta U = \hbar\omega_{ce}>0$, while the translational kinetic energy still is $\Delta T = \hbar k v_{\parallel}>0$. This would violate the conservation of energy, as it is not possible for an electron to emit a photon while simultaneously increasing its total energy. Consequently, for a plane electromagnetic wave, only the left-circularly polarized component can resonate with an electron moving opposite to $v_{\parallel}$ (or $\vec{B_0}$). 
%Based on momentum and angular momentum conservation analysis, we examine the case where $\vec{k}$ is oriented opposite to $\vec{v}_{\parallel}$ (or $\vec{B_0}$). 
%\begin{itemize}
%    \item \textbf{Left-hand circular polarization case}: When a cyclotron electron emits a photon with:
%
%    \begin{itemize}
%        \item Left-hand circular polarization ($\sigma^-$)
%        \item Momentum $-\hbar\vec{k}$
%        \item Angular momentum $+\hbar$
%    \end{itemize}
%    the resulting energy changes are:
%    \begin{align}
%        \Delta U &= -\hbar\omega_{ce} < 0 \quad \text{(decrease in internal energy)} \\
%        \Delta T &= +\hbar k v_{\parallel} > 0 \quad \text{(increase in translational energy)}
%    \end{align}
%
%    \item \textbf{Right-hand circular polarization case}: If the emitted photon had:
%    \begin{itemize}
%        \item Right-hand polarization ($\sigma^+$)
%        \item Momentum $-\hbar\vec{k}$
%        \item Angular momentum $-\hbar$
%    \end{itemize}
%    the energy changes would instead be:
%    \begin{align}
%        \Delta U &= +\hbar\omega_{ce} > 0 \quad \text{(increase in internal energy)} \\
%        \Delta T &= +\hbar k v_{\parallel} > 0 \quad \text{(increase in translational energy)}
%    \end{align}
%\end{itemize}

%The right-hand polarization case would violate energy conservation, as it requires the electron to \emph{emit} a photon while simultaneously \emph{gaining} total energy ($\Delta U + \Delta T > 0$). 
%Therefore, for a plane electromagnetic wave, only the left-circularly polarized component can resonantly interact with an electron moving antiparallel to $\vec{v}_{\parallel}$ while satisfying all conservation laws.
%Actually, this process could be also understand from classical view. Asfor ADR, where in our cause $\vec{k} \parallel \vec{v}$ as $v_z>c'$, the left hand polarization could reverse to right hand polarization in cyclotron electron frame, which could resonant with electron , and for NDR , where $v_z<c'$, the right hand polarization wave still keep the same polarization in cyclotron electron frame. but for $\vec{k}$ opposite to $v_z$ only left hand polarization wave have same rotation as electron ,no matter how fast the electron is , therefor only left hand polarization wave could resonant with electron in this case. 
%The resonance conditions can also be understood classically through polarization transformations in the electron's guiding center frame. For ADR ($\vec{k} \parallel \vec{v}_z$ with $v_z > c' \equiv \omega/|\vec{k}|$), an LHCP wave appears right-handed in the guiding frame, enabling resonance with electron  . Conversely, for NDR ($v_z < c'$), an RHCP wave maintains its polarization. When $\vec{k} \parallel -\vec{v}_z$, only LHCP waves maintain synchrony with the electron's gyration.
The resonant condition can also be understood from a classical perspective. In the case of ADR, where in our model $\vec{k} \parallel \vec{v}$ and $v_z > c' \equiv \omega/|\vec{k}|$, the LHCP wave appears as a right-hand polarized wave in the cyclotron electron's rest frame and can thus resonate with the electron. In contrast, for Normal Doppler Resonance , where $v_z < c'$, the RHCP wave maintains its polarization in the electron's rest frame. However, when $\vec{k}$ is anti-parallel to $\vec{v}$ (i.e., $\vec{k} \cdot \vec{v} < 0$), only the LHCP wave maintains the same rotational direction as the electron's cyclotron motion regardless of electron's parallel velocity. Therefore, in this configuration, only left-hand polarized waves can resonate with the electron.

This study also offers a  perspective on electron heating and current drive by electromagnetic (EM) waves. For example, during the NDR process, the ratio of the electron's internal energy gain from the EM wave to the total absorbed wave energy can be expressed as  
\(
\eta_H = \frac{m \omega_{ce}}{\omega},
\)
while the ratio contributing to the parallel kinetic energy is  
\(
\eta_T = \frac{\vec{k} \cdot \vec{v}}{\omega}.
\)
This effect may help optimize plasma heating and current drive efficiency. In the case of ADR, the electron's parallel kinetic energy can be converted into internal (gyro) energy. This mechanism may contribute to suppressing runaway electron energies, a phenomenon previously studied \cite{guo2018control,zhang2024self} but requiring further investigation in future fusion tokamak plasmas.

\section{Conclusion}\label{sec:Conclusion}
This paper presents a simple yet effective method for analyzing the resonant processes associated with the Normal Doppler Effect (NDE) and Anomalous Doppler Effect (ADE). By combining quantum theory with angular momentum conservation analysis, it is shown that the parameter \( m \) in the resonance condition  
\(
\omega = \vec{k} \cdot \vec{v} + m\omega_{ce}
\)
is directly associated with the angular momentum of the resonant wave. Numerical simulations based on the Volume Preserving Algorithm (VPA) further support the quantum results, confirming both the angular momentum correspondence of \( m \) and the energy transfer characteristics. In future work, the interaction between electron energy transformation and helicon waves will be investigated within a plasma environment, aiming to provide deeper insights into applications such as plasma heating and runaway electron suppression.
\section{Acknowledgements}
This work is supported by National Magnetic Confinement Fusion Energy Program of China under Contract No. 2019YFE03020001
%\addcontentsline{toc}{chapter}{Appendix A: Appendix section heading}
\section*{Appendix A:Classical Analysis of Anomalous Doppler Resonance}
Neglecting the static electric field and relativistic effect , we provide a brief derivation of the energy transformation process based on classical dynamical equations:
\begin{align}
m_e \frac{d\vec{v}_\parallel}{dt} &= -e(\vec{v}_\perp \times \vec{B}_\perp) \label{eq:19} \\
m_e \frac{d\vec{v}_\perp}{dt} &= -e(\vec{v}_\perp \times \vec{B}_0 + \vec{v}_\parallel \times \vec{B}_\perp + \vec{v}_\perp \times \vec{B}_0) \label{eq:20}
\end{align}
